Design and Implementation of Probabilistic Programming Language Anglican

### 2017-02-17 paper/paper.tex
update
minor

\subsection*{Contributions} This paper brings the following
major contributions:
\begin{itemize}
    \item Design and implementation of a probabilistic programming language
        Anglican involving tight bilateral integration with a general-purpose
        programming language.
    \item Techniques for efficient and compact implementation of inference
        algorithms, such as representation of inference results as a lazy
        sequence of samples and a novel scheme for addressing of checkpoints.
    \item Anglican's novel representation of random processes 
            from statistics and machine learning, such as Beta-Binomial 
            process. The representation is
            \emph{stateless} and seamlessly integrates into pure functional 
            computation. This contrasts with typical \emph{stateful}
            implementations of these processes in other probabilistic 
            programming languages.
            These constructs allowrepresents these p
%    \item Specification of probabilistic forms in Anglican which facilitates 
%        separation of probabilistic programs from inference algorithms.
\end{itemize}

@@

Probabilistic programming languages are implemented as
interpreters~\cite{GMR+08,MSP14}, embedded
languages~\cite{P09,SGG15,GSX+16}, compiled
languages~\cite{MMR+07,SDT14}, as through source-to-source
transformation with augmentation~\cite{WSG11,GS15}. Each of
these methods emphasizes different priorities in language
design, such as computational efficiency, integration with an
existing development environment, ease of implementation, or
efficiency of inference. Anglican implemented using a
combination of embedding and source-to-source transformational
compilation to combine advantages of both approaches.  CPS
transformation, employed by Anglican compiler, is also used in
WebPPL~\cite{GS15} and facilitates clean separation between
probabilistic programs and inference algorithms and diversity of
applicable inference algorithms.

Probabilistic programming languages are often implemented as
pure functional languages~\cite{GMR+08,MSP14,SGG15}. This is
because reasoning about probabilistic programs and distributions
defined by the programs, as well as implementation of certain
inference algorithms~\cite{MYM+15}, is easier in pure functional
setting. Anglican is also implemented as a language without
side-effects.  However, there are also languages which follow
imperative paradigm and support mutable state
directly~\cite{SDT14,GSX+16}, or through probabilistic
primitives~\cite{GMR+08,MSP14}. Anglican introduces a functional
alternative to stateful random primitives in the form of random
processes.

>>>>

\subsection*{Contributions} This paper brings the following
major contributions:
\begin{itemize}
    \item Design and implementation of a probabilistic programming language
        Anglican involving tight bilateral integration with a general-purpose
        programming language.
    \item Techniques for efficient and compact implementation of inference
        algorithms, such as representation of inference results as a lazy
        sequence of samples and a novel scheme for addressing of checkpoints.
    \item Anglican's novel representation of random processes 
            from statistics and machine learning, such as Beta-Binomial 
            process. The representation is
            \emph{stateless} and seamlessly integrates into pure functional 
            computation. This contrasts with typical \emph{stateful}
            implementations of these processes in other probabilistic 
            programming languages.
%    \item Specification of probabilistic forms in Anglican which facilitates 
%        separation of probabilistic programs from inference algorithms.
\end{itemize}

@@

Probabilistic programming languages are implemented as
interpreters~\cite{GMR+08,MSP14}, embedded
languages~\cite{P09,SGG15,GSX+16}, and compiled
languages~\cite{MMR+07,SDT14}, and also through source-to-source
transformation with augmentation~\cite{WSG11,GS15}. Each of
these methods emphasizes different priorities in language
design, such as computational efficiency, integration with an
existing development environment, ease of implementation, or
efficiency of inference. Anglican is implemented using a
combination of embedding and source-to-source transformational
compilation to combine advantages of both approaches.  CPS
transformation, employed by Anglican compiler, is also used in
WebPPL~\cite{GS15} and facilitates clean separation between
probabilistic programs and inference algorithms and diversity of
applicable inference algorithms.

Probabilistic programming languages are often implemented as
pure functional languages~\cite{GMR+08,MSP14,SGG15} in
the sense that they do not allow mutable states. This is
because reasoning about probabilistic programs and distributions
defined by the programs, as well as implementation of certain
inference algorithms~\cite{MYM+15}, is easier in such pure functional
setting. Anglican is also implemented as a language without
mutable states.  However, there are also languages which follow
imperative paradigm and support mutable state
directly~\cite{SDT14,GSX+16}, or through probabilistic
primitives~\cite{GMR+08,MSP14}. Anglican introduces a functional
alternative to stateful random primitives in the form of random
processes.
